Climate indicators

### 2023-02-24 DRAFT_climate_indicators.Rmd
microbenchmark

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
dir <- substr(getwd(), 1,2)

@@

For this example I want to calculate the mean temperature over the summer. I therefore need to aggregate over time. Here is how.

@@

```{r}
x_summerMean <- x_aug %>%
  filter(band ==1) %>% # for our real data we don't need this filter
  st_apply(1:2, mean)

dim(x_summerMean)
```

@@

*Method 2:*

We can also use the aggregate function to produce a data cube with lower temporal resolution, for example per year:

```{r}
x_summerMean_v2 <- x_aug %>%
  aggregate(by = "year", mean)
```

Actually, this second method is better because we don't need to filter out additional dimensions, and it retains the attribute name (the first method replaced it with the function name *mean*).

@@

Firt, let me cut the _band_ dimension.
```{r}
temp <- x_summerMean_v2 %>%
  filter(band==1) %>%
  adrop(drop=4) %>%
  aperm(c(2,3,1))
dim(temp)
```

>>>>

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
library(lubridate)
library(ggpubr)
library(microbenchmark)
dir <- substr(getwd(), 1,2)

@@

For this example I want to calculate the mean temperature over the summer. I therefore need to aggregate over time. There are two ways.

@@

```{r}
microOut <- microbenchmark(
st_apply =  x_aug %>%
  filter(band ==1) %>% # for our real data we don't need this filter
  st_apply(1:2, mean, rename=F),
aggregate = x_aug %>%
  aggregate(by = "year", mean),
times=30
)
```

@@

```{r, fig.cap="Comparing computation tome for two spatial aggregation functions."}
autoplot(microOut)
```

`st_apply` is considerably faster, even wit the filtering step, but `aggregate` has the advantage of that it retains the time dimension. 
I think I will try to use `aggregate`, and it it takes too long on the real data, then I can come back to this point, and perhaps switch to st_apply.

```{r}
x_summerMean <-   x_aug %>%
  filter(band ==1) %>% 
  st_apply(1:2, mean, rename=F)
```

@@

First, let me cut the _band_ dimension.
```{r}
temp <- x_summerMean_v2 %>%
  filter(band==1) %>%
  adrop(drop=4) %>%
  aperm(c(2,3,1))
dim(temp)
```
